---
title: "Getting Started with NanoQuRe"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with NanoQuRe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  message  = FALSE,
  warning  = FALSE
)
```

## Overview

NanoQuRe provides quality control tools for Oxford Nanopore sequencing data.
Starting from a sequencing summary file produced by ONT basecallers (Guppy or
Dorado), NanoQuRe generates interactive plots and summary tables — and can bundle
everything into a self-contained HTML report with a single function call.

## Installation

```{r install, eval = FALSE}
# install.packages("devtools")
devtools::install_github("All-ice0/NanoQuRe")
```

## Input data

NanoQuRe expects a data frame with the columns produced by ONT basecallers.
The minimum required columns are:

| Column | Type | Description |
|---|---|---|
| `sample_id` | character | Sample identifier |
| `run_id` | character | Run identifier |
| `channel` | integer | Pore channel number |
| `start_time` | numeric | Read start time in seconds |
| `duration` | numeric | Read duration in seconds |
| `sequence_length_template` | numeric | Read length in bases |
| `mean_qscore_template` | numeric | Mean Q score of the read |
| `passes_filtering` | logical | Whether the read passed basecaller filters |

Loading your own file is straightforward:

```{r load-real, eval = FALSE}
seq_summary <- read.csv("sequencing_summary.txt", sep = "\t")

# passes_filtering comes in as a string from some basecallers — coerce if needed
seq_summary$passes_filtering <- as.logical(seq_summary$passes_filtering)
```

For this vignette we use the example dataset bundled with the package:

```{r load-data}
library(NanoQuRe)

data(sample_data)
```

## Generating a full QC report

The main entry point is `generate_report()`. It inspects how many unique
`sample_id` values are in your data and automatically picks the right template
— single-sample or multi-sample — then renders a self-contained HTML file.

```{r generate-report, eval = FALSE}
generate_report(sample_data, output_file = "my_QC_report.html")
```

The rendered file will open in your browser automatically and contains all
plots and summary tables covering read quality, yield, throughput, and pore
activity.

## Multi-sample reports

If you have multiple samples, bind their data frames together before calling
`generate_report()`. The function detects more than one unique `sample_id`
and switches to the multi-sample template automatically:

```{r multi-sample, eval = FALSE}
generate_report(sample_data_A, sample_data_B, output_file = "multi_QC_report.html")
```

## What the report contains

The report bundles the output of all individual NanoQuRe functions.
If you want to explore or customise any single plot, the
[Using Individual Functions](individual-functions.html) vignette covers each
one in detail.

| Function | What it shows |
|---|---|
| `sequencing_stats()` | Run-level summary table |
| `quality_stats()` | Quality metrics table (N50, mean Q, longest read) |
| `plot_cumulative_yield()` | Cumulative bases over time, pass vs fail |
| `plot_seq_throughput()` | Yield per hour over time |
| `plot_average_speed()` | Translocation speed (bp/s) over time |
| `plot_read_lengths()` | Read length distribution with mean and N50 lines |
| `plot_quality_distribution()` | Q score distribution with pass/fail split |
| `plot_active_channels()` | Channel activity decay over time |
| `pore_activity_heatmap()` | Per-channel activity heatmap on flowcell layout |

## Session info

```{r session-info}
sessionInfo()
```