---
title: "Using Individual Functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Individual Functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  message  = FALSE,
  warning  = FALSE
)
```

## Overview

While `generate_report()` bundles everything into one HTML file, every plot
and table in NanoQuRe is also available as a standalone function. This is
useful when you want to embed a specific plot in your own R Markdown report,
tweak parameters, or inspect a single metric interactively.

```{r load-data}
library(NanoQuRe)

data(sample_data)
```

---

## Summary tables

### `sequencing_stats()`

Returns a one-row summary of the run: duration, total bases sequenced,
pass rate, and average translocation speed.

```{r sequencing-stats}
sequencing_stats(sample_data)
```

### `quality_stats()`

Returns key quality metrics: total and passed read counts, N50, mean Q score,
longest read, and mean length of passed reads.

```{r quality-stats}
quality_stats(sample_data)
```

---

## Yield plots

### `plot_cumulative_yield()`

Shows how total sequenced bases accumulate over the run, split by pass/fail
status. A flattening curve indicates the run is winding down.

```{r cumulative-yield}
plot_cumulative_yield(sample_data)
```

### `plot_seq_throughput()`

Shows yield per hour rather than cumulative yield — useful for spotting
periods of low pore activity or blockages during the run.

```{r seq-throughput}
plot_seq_throughput(sample_data)
```

---

## Speed

### `plot_average_speed()`

Shows the average translocation speed of DNA through the pore in bases per
second, split by pass/fail. A drop in speed can indicate pore aging or
buffer issues.

```{r average-speed}
plot_average_speed(sample_data)
```

---

## Read length distribution

### `plot_read_lengths()`

Plots the read length distribution with vertical lines for the mean read
length and the N50 value. The `upper_limit` parameter controls the x axis
maximum — useful when a small number of very long reads would compress
the rest of the distribution.

```{r read-lengths}
# default upper limit is 4000 bp
plot_read_lengths(sample_data)
```

```{r read-lengths-custom, eval = FALSE}
# zoom out to see the full long-read tail
plot_read_lengths(sample_data, upper_limit = 20000)
```

---

## Quality distribution

### `plot_quality_distribution()`

Plots the Q score distribution split by pass/fail. Two vertical lines are
shown: the basecaller's own pass/fail threshold (derived from the data) and
a user-defined cutoff for downstream analysis.

```{r quality-distribution}
# default user-defined cutoff is Q7
plot_quality_distribution(sample_data, qscore_cutoff = 7)
```

---

## Channel activity

### `plot_active_channels()`

Shows how many pore channels remain active over the course of the run.
A steep early drop suggests many pores became blocked or inactive quickly.

```{r active-channels}
plot_active_channels(sample_data)
```

### `pore_activity_heatmap()`

Shows two complementary views of pore activity: a time × channel heatmap
and a spatial heatmap overlaid on the flowcell layout. Darker colour means
more bases sequenced through that channel.

```{r pore-heatmap}
pore_activity_heatmap(sample_data, platform = "minion")
```

---

## N50 directly

### `calculate_n50()`

If you just need the N50 value without any plot:

```{r n50}
calculate_n50(sample_data)
```

---

## Session info

```{r session-info}
sessionInfo()
```